---
title: "P8105 HW1"
author: "Jennifer Lee (UNI: jl2361)"
output: github_document
---

## Problem 1

### Description of the dataset
We will load the tidyverse.

```{r setup}
library(tidyverse)
```

We will load in the `penguins` **dataset**. Here is a view of the dataset and a brief summary.

```{r}
data("penguins", package = "palmerpenguins")
head(penguins)
summary(penguins)
```

There are data from 3 species of penguins of approximately equal sex distribution from 3 different islands. There are **8 total variables** in this dataset: `species`, `island`, `bill_length__mm`, `bill_depth_mm`, `flipper_length_mm`, `body_mass_g`, `sex`, and `year`. 

There are **`r nrow(penguins)` rows** and **`r ncol(penguins)` columns**. 

The **mean flipper length** is `r mean(pull(penguins, flipper_length_mm), na.rm = TRUE)` mm. 

### Scatter plot
Next, this is a **scatter plot** with `flipper_length_mm` (y) vs and `bill_length_mm` (x) with points colored using the `species` variable.

```{r}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()
```

Finally, we will **export** this scatter plot to the project directory as a PDF file.

```{r}
ggsave("scatterplot.pdf", height = 4, width = 6)
```

## Problem 2 
First, we will create our **dataframe** comprised of:

* A random sample of size 10 from a standard Normal distribution
* A logical vector indicating whether elements of the sample are greater than 0
* A character vector of length 10
* A factor vector of length 10, with 3 different factor “levels”

```{r}
df = tibble(
  random_sample = rnorm(10),
  vec_logic = random_sample > 0,
  vec_char = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
  vec_factor = factor(c("high", "medium", "low", "high", "medium", "low", "high", "medium", "low", "high"))
)

print(df)
```

Next, we will try to take the **mean** of each variable in the dataframe. We are able to obtain the mean of `random_sample` since it is a numeric vector. We are also able to obtain the mean of `vec_logic` where `TRUE` is coerced to a numeric of 1 and `FALSE` is coerced to a numeric of 0. We are unable to obtain the means of `vec_char` or `vec_factor`.
```{r}
# mean of random_sample variable- this works
mean(pull(df, random_sample))

# mean of vec_logic variable- this works
mean(pull(df, vec_logic))

# mean of vec_char variable- this does not work
mean(pull(df, vec_char))

# mean of vec_factor variable- this does not work
mean(pull(df, vec_factor))
```

We will now explicitly **convert** the logical (`vec_logic`), character (`vec_char`), and factor (`vec_factor`) variables using the `as.numeric` function. This function converts `vec_logic` and `vec_factor` to the numeric type. Again, we are able to obtain the mean of `vec_logic` where `TRUE` is coerced to a numeric of 1 and `FALSE` is coerced to a numeric of 0. If we were to create a new variable for `vec_factor` converted to a numeric vector using the `as.numeric` function, we would then be able to obtain the mean of this new variable. 

The `as.numeric` function is, however, unable to convert `vec_char` to the numeric type, so we are not able to obtain its mean. 
```{r eval = FALSE}
as.numeric(pull(df, vec_logic)) 
as.numeric(pull(df, vec_char)) 
as.numeric(pull(df, vec_factor))

# mean of vec_logic variable- this works
mean(pull(df, vec_logic))

# mean of vec_char variable- this does not work 
mean(pull(df, vec_char))

# mean of vec_factor variable- this does not work unless we create a new variable for vec_factor converted to a numeric type using the as.numeric function
mean(pull(df, vec_factor))
```